Run AwAI: StarCraftII Adversary Avoidance Using Neural Networks and Genetic Algorithms

### 2018-06-04 doc/run_awai.tex
added controller writeup

>>>>

The Pysc2 interface allowed us to create an agent to interact with the Star
Craft II game.  The agent class contains a \texttt{step()} function that is called
several times per second of game time. This meant that all of the actions the
agent took had to be in this step function, and thus this is where we would
integrate our neural net.  We decided to use the mini game map DefeatRoaches,
which is included with Pysc2 as a base for our simulation environment. By
heavily relying on a preexisting map and mini game, we could focus more of our
time and resources on developing the genetic algorithm and neural network
implementation.

To make our controller implementation as modular as possible, we decided to
provide a dictionary of inputs and receive a dictionary of outputs. The inputs
would be the x and y coordinates of the agent’s units, as well as the x and y
coordinates of the enemy units. The inputs had to be created and assembled into
a dictionary in the step function.  In order to normalize the inputs, we divided
the coordinate value by the maximum map size in the corresponding axis.  To
obtain the values for the inputs, we created a series of methods for our agent
class.  These methods include \texttt{getCurrentLocation()} and
\texttt{getCurrentEnemeyLocation}().  We also created methods to move the friendly units,
including \texttt{moveToLocation()} and setTargetDestination.  These functions could be
used together to set a target destination coordinate and cause the friendly
units to move toward that target.

One problem we encountered that we had to correct was determining when units
arrived at their target destination.  We solved this by comparing the target
destination and the units current location every time \texttt{step()} is called.  Because
the DefeatRoaches mini game uses multiple units (both friendly and hostile), the
unit location coordinates are averages of all selected units locations.  This
means that often units will never actually arrive at their precise destination,
just get very close to it.  To mitigate this issue, we decided to consider units
within a certain radius of their target destination as at their destination.
Once this condition is met, the \texttt{moveToLocation()} function returns a status of
arrived at target.

Whenever \texttt{step()} is called by Pysc2, the first thing we do is check the
observable action space. If friendly unit movement is not an option, this means
the unit(s) are not selected.  If units are not selected, this means it is the
first time the \texttt{step()} function is being called in the simulation episode.  In
this case, in addition to selecting the units, we also reset various counters
(such as steps survived).  During the first call of \texttt{step()}, we also unpickle a
neural network file.

In the case where \texttt{step()} is called and units are already selected, the
\texttt{moveToLocation()} function is called.  Initially, we set the target destination
to the enemy location.  We have our units move directly to the hostile units in
order to engage them in combat, and this initiate a pursuit from the enemy.  We
felt this would be the most direct way to initiate a pursuit.  We investigated
running two Pysc2 agents against each other, however Pysc2 currently only
support the simulation of one custom agent at a time. Once the friendly units
arrive at their destination, the \texttt{moveToLocation()} function returns an arrived at
target status to the calling \texttt{step()} function.

When the \texttt{step()} function receives an arrived at target status, a new target
destination is set.  To generate a new target destination, the relevant aspects
of the current state (current location of friendly and enemy units) are captured
and fed into the neural network.  The neural network returns a dictionary
containing the action space and a value between -1 and 1 for each action label
key.  The action with the highest number will be chosen, and then taken.

Our action space is movements that the friendly units can take.  We decided to
quantize the action space into discrete movements, as opposed to a continuous
x-y coordinate plane. By only allowing our units to move in the 4 cardinal
directions, diagonals, or stay put, we hoped our first implementation of the
neural network and genetic algorithm would have an easier time converging.  In
the future, we could expand the output action space to a continuous x and y
coordinate plane, as opposed to discrete directions. Although a larger action
space can often require more iterations of the genetic algorithm to converge, as
the problem complexity is much higher.

To evaluate the fitness of the neural network, we count the number of steps that
the units survived for.  One problem we encountered and solved, was the initial
series of steps taken before engaging the enemy units. For example, the
DefeatRoaches map spawns the enemy and friendly units on opposite sides of the
map, in a random selection of several predefined start locations.  However, the
distance between these start locations can vary between simulation episode.
This means our neural network fitness (number of steps survived) could be
slightly affected by an irrelevant condition.  To mitigate this issue, we
subtract the number of steps it takes until the friendly units reach their first
destination (the location of the enemy units) and start their engagement. By
only counting steps after enemy engagement has begun, our fitness function is
more reflective of survival time.

In order for our evolutionary genetic algorithm to evolve, the end fitness of a
network must be returned from the Pysc2 simulation.  Our simulation runs a
configurable (up to 55) episodes per individual (neural network) and averages
the fitness for all of those simulation episodes. We played with the number of
simulation episodes per simulation test, exploring the tradeoff between
simulation runtime and precision.  In the future, we could explore modifying our
fitness function to incorporate more meaning from having multiple simulation
episodes.  One future idea we had was to see if the group of individual survival
times were skewed above or below the mean.

The average fitness for the individual is then saved in a pickle file, along
with the fitness of the top (most fit) individual.  Because the Pysc2 simulation
is invoked from the genetic algorithm code, the algorithm can then unpickle the
file.  After all individuals in a generation are tested and their fitnesses
collected, the genetic algorithm can select the most fit individuals (neural
networks) for breeding.
